## 2.2 Hypothesis Testing

```{r hyp-setup, include=FALSE}
library(tidyverse)
library(knitr)
load("data/bluesky_raw.RData")
```

**Research question.** *Are posts from human accounts more likely to receive engagement than posts from automated (bot) accounts?*

Section 2.1 showed that bots produce over half of the English cybersecurity posts. This test asks whether that volume translates into attention from other users.

We use the English posts **before** bot removal, applying the same duplicate, language and length filters as the cleaned data, and label each post by account type.

```{r hyp-data}
# Same bot list as scripts/Clean.R (13 flagged handles, 4 manually kept as real accounts)
flagged <- unique(posts_raw$author_handle[str_detect(posts_raw$author_handle, "bot")])
kept    <- c("ebottcher.bsky.social", "repeatablerobot.bsky.social",
             "adybot.bsky.social", "progressiverobot.bsky.social")
bots    <- setdiff(flagged, kept)

hyp <- posts_raw %>%
  distinct(uri, .keep_all = TRUE) %>%
  filter(map_lgl(langs, ~ "en" %in% unlist(.x))) %>%
  mutate(n_words      = str_count(str_squish(text), "\\S+"),
         account_type = if_else(author_handle %in% bots, "Bot", "Human"),
         engaged      = if_else(like_count + repost_count + reply_count > 0,
                                "Engaged", "No engagement")) %>%
  filter(n_words >= 5)

tab <- table(hyp$account_type, hyp$engaged)
tab
round(prop.table(tab, margin = 1), 3)      # proportion engaged within each group
```

```{r hyp-plot, fig.height=3, fig.width=5, fig.cap="Share of posts receiving any engagement, by account type"}
hyp %>%
  count(account_type, engaged) %>%
  group_by(account_type) %>% mutate(prop = n / sum(n)) %>%
  filter(engaged == "Engaged") %>%
  ggplot(aes(account_type, prop)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(x = NULL, y = "Posts with >= 1 like, repost or reply")
```

**Hypotheses.**

- H0: account type and receiving engagement are independent.
- H1: account type and receiving engagement are associated.

**Why this test.** Both variables are categorical (bot/human; engaged/not), so a chi-squared test of independence is appropriate. Engagement counts are heavily skewed (the median post receives 0 likes), so comparing means with a t-test would be misleading; converting to "any engagement" avoids this. The test requires expected counts of at least 5 in every cell, checked below.

```{r hyp-test}
test <- chisq.test(tab)
test$expected      # all should be >= 5
test
```

**Result and interpretation.** The chi-squared test of independence gave 	$\chi^2 = 547.67$ with 1 degree of freedom and $p < 2.2 \times 10^{-16}$, which is far below $\alpha = 0.05$, so we reject H0 and conclude that account type and receiving engagement are associated. The proportions show the direction of this association: 56.5% of human posts received at least one like, repost or reply, compared with only 14.5% of bot posts, making human posts roughly four times as likely to receive engagement. Combined with Section 2.1, where bots wrote 53% of English cybersecurity posts, this suggests that automated accounts dominate the volume of the conversation while users mostly direct their attention to human accounts. This is an association rather than a causal effect, since engagement also depends on factors such as follower counts and posting time that the test does not control for.

**Limitations.** Posts are not fully independent: a few bot accounts produced many posts each, so the test treats repeated posts from one account as separate observations. Engagement also depends on follower counts and posting time, which this test does not control for, so an association does not show that being a bot *causes* lower engagement.
