## 2.3 Text/Content Analysis and Visualisation

**Research question:** What are the most frequently discussed topics in cybersecurity-related posts on Bluesky?

**Cleaning and representation.** We start from the 1,293 cleaned posts (Section 2.1). Posts with empty or missing text are dropped, and URLs are removed. Each post is then split into lower-case word tokens, with punctuation removed. Standard English stop words are removed, along with the two search keywords ("cybersecurity", "infosec", which appear in almost every post) and URL fragments ("https", "www", "amp"). Pure numbers are also removed. The result is a one-word-per-row table that we use for the frequency summaries here. For clustering (Section 2.4), we convert it into a TF-IDF weighted document-term matrix.

```{r text-setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(knitr)

posts <- readRDS("data/posts_clean.rds")
```

```{r text-cleaning}
custom_stop <- tibble(
  word = c("cybersecurity", "infosec",
           "https", "http", "www", "amp")
)

words <- posts %>%
  select(id, text) %>%
  filter(!is.na(text), text != "") %>%
  mutate(text = str_remove_all(text, "https?://\\S+")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") %>%
  anti_join(custom_stop, by = "word") %>%
  filter(!str_detect(word, "^[0-9]+$"))

word_frequency <- words %>%
  count(word, sort = TRUE)
```

```{r word-frequency-plot, fig.width=7, fig.height=4.5, fig.cap="The 20 most frequent words in cybersecurity posts (search keywords and stop words removed)"}
word_frequency %>%
  slice_head(n = 20) %>%
  ggplot(aes(x = n, y = reorder(word, n))) +
  geom_col(fill = "#6B5B95") +
  labs(x = "Frequency", y = NULL) +
  theme_minimal()
```

```{r word-authors-plot, fig.width=7, fig.height=4, fig.cap="Number of distinct authors using each of the 15 most frequent words"}
word_author_counts <- words %>%
  distinct(id, word) %>%
  left_join(posts %>% select(id, author), by = "id") %>%
  group_by(word) %>%
  summarise(authors = n_distinct(author), .groups = "drop")

word_frequency %>%
  inner_join(word_author_counts, by = "word") %>%
  slice_max(n, n = 15) %>%
  ggplot(aes(x = authors, y = reorder(word, authors))) +
  geom_col(fill = "#6B5B95") +
  labs(x = "Distinct authors", y = NULL) +
  theme_minimal()
```
```{r author-counts, include=FALSE}
n_auth <- function(w) word_author_counts$authors[word_author_counts$word == w]
```

**Interpretation.** The frequency chart shows that AI, ransomware, threat intelligence and vulnerabilities are the most prominent topics. However, counting authors changes the picture. "AI" was used by `r n_auth("ai")` different authors, so it is a topic discussed across the community. "Ransomware" came from `r n_auth("ransomware")` authors, and "CTI" and "threatintelligence" from only `r n_auth("cti")` and `r n_auth("threatintelligence")`. A few highly active accounts can therefore make a niche topic look dominant in raw word counts. This matches the automated feeds noted in Sections 2.1 and 2.2, such as cti.fyi. Repeated meta-words like "post", "blog" and "title" also suggest templated posts from news feeds. Because frequency alone mixes broad discussion with high-volume posting, Section 2.4 uses clustering to separate the main types of post.
