## 2.4 Clustering

**Research question:** What groups of similar topics can be identified in cybersecurity-related Bluesky posts?

**Decisions.**

- *Representation:* each post is represented as a TF-IDF weighted document-term matrix, built from the cleaned tokens in Section 2.3. TF-IDF gives less weight to words that appear in many posts and more weight to words that distinguish a post. To reduce noise and sparsity, we keep only the 500 most frequent terms.
- *Distance:* posts differ greatly in length, so we use cosine distance, which compares the direction of word profiles rather than their magnitude. Each post vector is scaled to unit length. After that scaling, the Euclidean distance used by k-means ranks posts the same way as cosine distance.
- *Number of clusters:* k is chosen with the elbow method. We plot the total within-cluster sum of squares for k = 2 to 10 and pick the point where adding more clusters stops giving large improvements.

```{r clustering-setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(knitr)

posts <- readRDS("data/posts_clean.rds")
```

```{r tfidf}
custom_stop <- tibble(
  word = c("cybersecurity", "infosec",
           "https", "http", "www", "amp")
)

cluster_words <- posts %>%
  select(id, text) %>%
  filter(!is.na(text), text != "") %>%
  mutate(text = str_remove_all(text, "https?://\\S+")) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words, by = "word") %>%
  anti_join(custom_stop, by = "word") %>%
  filter(!str_detect(word, "^[0-9]+$"))

tfidf_data <- cluster_words %>%
  count(id, word) %>%
  bind_tf_idf(word, id, n)

# Keep the 500 most frequent words to reduce noise
top_terms <- cluster_words %>%
  count(word, sort = TRUE) %>%
  slice_head(n = 500) %>%
  pull(word)

# TF-IDF document-term matrix (posts in rows, terms in columns)
tfidf_matrix <- tfidf_data %>%
  filter(word %in% top_terms) %>%
  select(id, word, tf_idf) %>%
  pivot_wider(names_from = word, values_from = tf_idf, values_fill = 0)

post_ids <- tfidf_matrix$id
X <- as.matrix(tfidf_matrix %>% select(-id))

# Remove posts with no retained terms
keep <- rowSums(X^2) > 0
X <- X[keep, , drop = FALSE]
post_ids <- post_ids[keep]

# Unit-length rows, so Euclidean distance behaves like cosine distance
X <- X / sqrt(rowSums(X^2))
dim(X)
```

```{r elbow, fig.width=6, fig.height=3.5, fig.cap="Elbow plot: total within-cluster sum of squares for k = 2 to 10"}
set.seed(123)
k_values <- 2:10
wss <- sapply(k_values, function(k) {
  kmeans(X, centers = k, nstart = 10, iter.max = 100)$tot.withinss
})
plot(k_values, wss, type = "b", pch = 19,
     xlab = "Number of clusters (k)",
     ylab = "Total within-cluster SS")
```

The elbow is gradual rather than sharp, but the curve starts to flatten at about k = 6, so we use six clusters.

```{r kmeans}
set.seed(123)
k <- 6
km <- kmeans(X, centers = k, nstart = 25, iter.max = 100)

# Relabel clusters by size, largest first
cluster_sizes <- sort(table(km$cluster), decreasing = TRUE)

cluster_order <- as.integer(names(cluster_sizes))

new_labels <- setNames(
  seq_along(cluster_order),
  cluster_order
)

cluster_assignments <- tibble(
  id = post_ids,
  cluster = factor(
    new_labels[as.character(km$cluster)],
    levels = 1:6
  )
)

cluster_terms <- tfidf_data %>%
  inner_join(cluster_assignments, by = "id") %>%
  group_by(cluster, word) %>%
  summarise(
    mean_tfidf = sum(tf_idf) /
      n_distinct(cluster_assignments$id[cluster_assignments$cluster == first(cluster)]),
    .groups = "drop"
  ) %>%
  group_by(cluster) %>%
  slice_max(mean_tfidf, n = 10) %>%
  arrange(cluster, desc(mean_tfidf))
```

```{r cluster-summary}
cluster_assignments %>%
  count(cluster, name = "posts") %>%
  mutate(share = scales::percent(posts / sum(posts), accuracy = 1)) %>%
  left_join(cluster_terms %>%
              group_by(cluster) %>%
              slice_head(n = 6) %>%
              summarise(top_terms = paste(word, collapse = ", ")),
            by = "cluster") %>%
  kable(caption = "Size and top TF-IDF terms of each cluster")
```

```{r cluster-mds, fig.width=6, fig.height=4.5, fig.cap="Posts projected into two dimensions with classical MDS on the cosine-based distances, coloured by cluster"}
mds <- cmdscale(dist(X), k = 2)
plot(mds, col = as.integer(cluster_assignments$cluster), pch = 19, cex = 0.5,
     xlab = "Dimension 1", ylab = "Dimension 2")
legend("topright", legend = paste("Cluster", 1:k), col = 1:k,
       pch = 19, cex = 0.7)
        
```

```{r cluster-terms-plot, fig.width=7, fig.height=6, fig.cap="Top words characterising each cluster (mean TF-IDF)"}
cluster_terms %>%
  group_by(cluster) %>%
  slice_max(mean_tfidf, n = 8) %>%
  ungroup() %>%
  ggplot(aes(x = mean_tfidf,
             y = reorder_within(word, mean_tfidf, cluster),
             fill = cluster)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~cluster, scales = "free_y", ncol = 2) +
  scale_y_reordered() +
  labs(x = "Mean TF-IDF", y = NULL) +
  theme_minimal()
```

```{r save-clusters, include=FALSE}
posts_clustered <- posts %>% left_join(cluster_assignments, by = "id")
saveRDS(posts_clustered, "data/posts_clustered.rds")
```

### Interpretation of Clusters

The K-means analysis identified six groups of cybersecurity-related posts, ranging from general discussions to specialised security alerts and employment opportunities.

**Cluster 1 — General Cybersecurity Discussion (780 posts):** This was the largest cluster, representing approximately 61.9% of clustered posts. Its most characteristic terms included *data*, *security*, *hacking*, *people* and *news*. The cluster reflects broad cybersecurity conversations rather than a single specialised topic.

**Cluster 2 — AI and Security News (191 posts):** This cluster was characterised by terms such as *AI*, *OpenAI*, *security* and *Google*. It captures discussions about artificial intelligence, emerging technologies and their implications for cybersecurity. This finding also supports the text analysis in Section 2.3, where AI was one of the most frequently discussed topics across a broad range of authors.

**Cluster 3 — Ransomware and Threat-Intelligence Alerts (190 posts):** Terms including *CTI*, *threatintelligence*, *ransom*, *post* and *title* dominated this cluster. These terms suggest structured threat-intelligence reporting and repeated ransomware-related alerts, rather than general conversation.

**Cluster 4 — Software Vulnerabilities and CVE Alerts (52 posts):** The distinctive vocabulary included *CVE*, *Chromium* and *Debian*. This cluster primarily concerns software vulnerabilities, security advisories and technical exploit information.

**Cluster 5 — German-Language Cyber Incidents (29 posts):** Words such as *Cyberangriff*, *Angriff*, *Einbruch* and *Datendiebstahl* indicate German-language reporting about cyberattacks, intrusions and data theft. This reveals a limitation of the English-language filtering process, as some non-English content remained in the dataset.

**Cluster 6 — Cybersecurity Employment Opportunities (19 posts):** This cluster contained terms such as *job*, *jobs*, *senior* and *educativ.net*, indicating cybersecurity recruitment advertisements and professional opportunities.

Overall, the clustering analysis demonstrates that cybersecurity-related posts on Bluesky are not homogeneous. Although general discussion accounts for the largest proportion of posts, the smaller clusters reveal specialised themes, including AI, vulnerability alerts, ransomware intelligence, cyber incidents and recruitment.

The results also illustrate how repeated content from a small number of accounts can influence the apparent prominence of particular topics.
**Limitations.** The elbow is not sharp, so other values of k could be argued for. K-means also depends on the random starting points, and we reduce this by using 25 starts and a fixed seed. Keeping only the 500 most frequent terms may merge rarer topics into the large general cluster.
