---
title: "Who Shapes Cybersecurity Conversation on Bluesky?"
subtitle: "Social Web Analytics (COMP3020): Group Project Report"
output:
  pdf_document:
    number_sections: false
geometry: margin=2cm
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.height = 3.5, fig.width = 6)
options(knitr.kable.NA = "–")
library(tidyverse)
library(knitr)
```

**Group 32**

\vspace{0.5em}
```{r members, echo=FALSE}
tibble(
  `Group member` = c("Yousuf Hossain Munna", "Seemrah Panwala", "Rishikanth Goud Nimmala"),
   SID            = c("20397249", "22175168", "22170637"), 
   Contribution   = c("40%", "30%", "30%"),
  `Main sections` = c("2.1 Data collection, 2.2 Hypothesis testing, Report assembly & Github",
                      "2.3 Text analysis, 2.4 Clustering",
                      "2.5 Network analysis, Poster")
) %>% kable (format = "latex", booktabs = TRUE,
            align = c("l", "l", "c", "p{7cm}"))

```
\vspace{1em}

**Code and data repository:** <https://github.com/Yousuf963/comp3020-group32>

# 1 Research Questions

This project examines who shapes cybersecurity conversation on Bluesky by analysing posts containing "cybersecurity" or "infosec" and a follow network surrounding a prominent cybersecurity account.

1. **Hypothesis (2.2):** Are posts from automated (bot) accounts less likely to receive engagement than posts from human managed accounts?

2. **Text (2.3):** What are the most frequently discussed topics in cybersecurity-related posts on Bluesky?

3. **Clustering (2.4):** What groups of similar topics can be identified in cybersecurity-related Bluesky posts?

4. **Network (2.5):** In the follow network around the most-followed active author in the cybersecurity discussion, which accounts are most central, what kind of accounts are they, and does this network belong to the cybersecurity community?

# 2 Analyses

```{r child = "01_data.Rmd"}
```

```{r child = "02_hypothesis.Rmd"}
```

```{r child = "03_text.Rmd", eval = file.exists("03_text.Rmd")}
```

```{r child = "04_clustering.Rmd", eval = file.exists("04_clustering.Rmd")}
```

```{r child = "05_network.Rmd", eval = file.exists("05_network.Rmd")}
```

# 3 Overall Findings

```{r child = "06_findings.Rmd", eval = file.exists("06_findings.Rmd")}
```
